\section*{Preface}
This master thesis was written as the paper for TMM4935 \-- Industrial ICT, Master's Thesis for the Department of of Mechanical and Industrial Engineering the Norwegian University of Science and Techonology (NTNU).

The development of ATO at NTNU has been in progress for multiple years and is now in the remote operation phase. Master students at the Department of Mechanical and Industrial Engineering has steadily worked on this project. In the fall of 2024 and spring of 2025 it was from Emilia Kozarevic, and I will continue her and NTNU's research on the topic.

Subject \& knowledge listed below has been used.
\begin{itemize}
    \item TBA4225 \-- Railway Engineering 2
    \item IT2810  \-- Web Development
    \item TDT4240 \-- Software Architecture
    \item TDT4120 \-- Algorithms and Data Structures
    \item TMM4540 \-- Industrial ICT Engineering Design and Materials, Specialization Project
\end{itemize}

For this master thesis, I would like to thank all my fellow students at Engineering \& ICT and the student union Hybrida. Hybrida and it's members have been crucial for me during my years studying at NTNU, and I owe so much to them all. If I were to do it again, I would have done it all the same as it has given me a lifetime of laughter, experience, friendships for life, a new me and love. Thank you so much. 

I would like to end the preface by thanking my fantastic and engaged supervisor Nils O.E. Olsson for the work of supervising me during this semester and the effort he has put in to help me write this paper. I also would like to thank Patrick Urassa from NTNU and Tomas Rosberg from VTI for the discussions and the insight in the project. Lastly I would like to thank my co-students Judith Røynstrand and Nathalia Thu for assisting me with testing of the final product.
