\section{Methodology}

\subsection{Literature and Knowledge acquisition}
The knowledge needed to carry out this thesis was acquired through an iterative process that combined systematic literature review with hands-on practical exploration. This approach, often referred to as an Iterative Literature and Practice-Driven Approach, is characterized by a continuous cycle where theory and practice inform each other rather than following a strict linear sequence [Wohlin et al. (2012)]. Instead of completing all literature review before beginning practical work, the two activities run in parallel, with new practical challenges prompting targeted literature searches and new theoretical insights shaping the direction of experiments. This makes it particularly well suited for applied engineering research where the problem space is not fully known at the outset and understanding develops gradually through doing [Hevner et al. (2004) Saunders et al. (2019) ].

The process began by identifying the core knowledge areas required before any development could start. This included understanding the operational context of remote train operation and Automatic Train Operation, the fundamentals of video streaming and relevant transport protocols, and the capabilities and structure of GStreamer as a framework. These areas formed the initial scope of the literature review and gave enough grounding to begin practical work.

Academic literature was gathered primarily through Google Scholar and IEEE Xplore, using search terms related to low-latency video streaming, remote train operation, GStreamer pipeline design, and network performance for real-time applications. Sources were selected based on relevance to the problem, publication quality, and recency where applicable. In addition to academic papers, official technical documentation was used extensively, particularly the GStreamer documentation and protocol specifications for RTP and UDP, which provided the practical detail that academic sources alone could not cover.

Once a sufficient theoretical foundation was in place, practical exploration began through hands-on work with GStreamer. Building and testing pipelines from the command line and through Python quickly revealed challenges that existing literature did not fully address, such as specific configuration trade-offs, unexpected latency behavior, and hardware compatibility issues. Each of these practical obstacles prompted a return to the literature, this time with more precise and targeted questions. New findings were then brought back into the development process, tested, and evaluated against real behaviour.

This cycle of reading, building, observing, and searching again repeated throughout the entire development phase. The result is that the theoretical foundation of this thesis did not precede the practical work, but grew alongside it, shaped by the specific demands of the problem at hand.



\subsubsection{Search Strategy and Databases}
A multi-platform search approach was utilized to retrieve a wide array of high-quality sources. The primary databases included:

\begin{itemize}
    \item Scopus: For retrieving peer-reviewed, impactful scientific articles.
    \item Google Scholar: For broader academic and institutional literature.
    \item GStreamer docs: For specifics on GStreamer development.
\end{itemize}



\subsubsection{Inclusion Criteria}
To ensure the study is based on the relevant and current information, specific criteria were applied to filter the search results. Given the rapid pace of technological change, a focus was placed on recent publications. Where modern papers have referenced to older papers, I have include some as to say and show where certain numbers are coming from and how they have influenced results.

\begin{table}[h]
    \centering
    \caption{Inclusion Criteria for Literature Review}
    \label{tab:inclusion_criteria}
    \begin{tabular}{|c|p{8cm}|}
        \hline
        ID & Inclusion Criteria (IC) \\
        \hline
        IC1 & Publication Date: Between 2020-2026 \\
        \hline
        IC2 & Language: Written in English \\
        \hline
        IC3 & Document Type: Primarily "Article", "Conference Paper" or "Thesis"\\
        \hline
    \end{tabular}
\end{table}


\subsubsection{Search Queries and Refinement}




Several digital tools as mentioned below were used for this paper. The main purpose of the usage was to enhance clarity, assisting in latex, and in general increase quality. Specifics include writing reference list in correct format. Structuring sentences and paragraphs. Checking for typos and grammatical errors.

\begin{itemize}
    \item OpenAI's ChatGPT: Employed as a helpful resource for LaTeX formatting suggestions, generating structured content (tables and lists), translating between Norwegian and English, and reviewing text for synonyms and restructuring ideas to ensure arguments were effectively communicated.
    \item Google Gemini: Used for similar fields as ChatGPT, but also for general proofreading, and refining sentence structure and tone to maintain a high standard of academic writing.
    \item Anthropic's Claude.ai for discussion and development of script syntax.
    \item Visual Studio Code: was used as the main LaTeX editor because of its powerful extensions for LaTeX support, syntax highlighting, and version control integration.
\end{itemize}

\subsection{Validity}

\subsection{Reliability}

\subsection{Research Design}
This thesis follows an applied research approach, combining Design Science Research (DSR) and experimental research to address both the development and evaluation aspects of the work. This combination is well suited for engineering projects where the goal is to produce a functional artifact while also generating knowledge through systematic testing and measurement.

The DSR component governs the design and development of the video streaming pipeline. Following the principles outlined by Hevner et al. [source], the research contribution is embedded in the artifact itself, meaning the pipeline design, the architectural decisions made, and the reasoning behind them all form part of the research output. The problem is clearly grounded in practice, the solution is built and refined iteratively, and the artifact is evaluated against the requirements defined in the previous chapter.

The experimental research component covers the testing and evaluation phase. Drawing on the framework described by Wohlin et al. [source], the pipeline is tested across a range of configurations in a controlled setting. Variables such as codec choice, payload settings, and network conditions are systematically varied, and end-to-end latency is measured and compared across configurations. This allows conclusions to be drawn about which configurations perform best and what trade-offs exist between latency and video quality.

Together these two approaches reflect the dual nature of the thesis: it is both a development project and a measurement study, and the methodology is designed to support both goals.
\subsection{Development}
\subsubsection{GStreamer Pipeline Development}
\subsection{Experimental Methodology}
\subsubsection{Latency Measurement}

