\section{Conclusion}

Conclusion needs to be drawn up to RQ and try to give a little closure of what the paper discussed. 
\begin{enumerate}
    \item RQ1: How is latency evaluated for railway and other industries? 
    \item RQ2: What parameters influence latency in remote train operation systems?
    \item RQ3: How can we test different video streaming protocols and their impact on real-time performance of remote train operation systems?
\end{enumerate}

This study explored how latency can be measured and evaluated for remote train operation systems, and how acceptable performance thresholds can be established. By reviewing research across railway, car, drone and crane control, and by analysing human responses alongside technical measurements, the study provides a clearer picture of what aspects of latency matter most and how they should be assessed.

\subsection{Evaluation of latency}
The findings show that latency evaluation in train operation might need different methods than in other industries. Remote operation research on cars, drones and cranes often have different tests requiring movements, such as driving slalom or avoiding hazards, that train operation do not acquire. Making many of the latency's found in the research of other industries invalid or irrelevant. 
Still there is many ways RTO can learn from these industries. By duplicating setups, experiements format, attributes considered when latency is evaluated can be replicated for RTO research. By dividing latency into one-way delay, round-trip time and glass-to-glass delay, and relies on synchronized timestamping to obtain accurate end-to-end measurements.


\subsection{Important aspects of latency}

\subsection{Video streaming tools}
There are many varients of tools and protocls to complete a video stream. 

Just as we have observed how different video streaming tools has been used, upgraded and replaced, we must anticipate that today's tools also will be upgraded. Not to mention the wast spectrum of different models, businesses and products on the market able to provide the video streaming service. 

As businesses try to improve their products, 


As discussed prievously, there is a lot of different ways to set up a remote control operation. You can choose from many different protocols, specialised hardware, or opt for a developed solution from an outside business. 
And as time grows, more option are destined to also arrive as the market are not yet satifised with todays level. If we are to achive the requiremtns set up by 3GPP for remote control under GoA~3 GoA~4 \cite{3gppts22289}, new protocols, hardware or a different way to do to remote operation is necessary. However, as of now, these requirment, should as mentioned in the paper be used as target values. When performing latency test